---
title: "PA1_template"
author: "VikingPrincess"
date: "Sunday, June 14, 2015"
output: html_document
---

First, let's load the data:

```{r, echo=TRUE}
data <- read.csv("activity.csv")
```
That is assuming we are in the working directory that contains a file "activity.csv".
To the task:

# What is mean total number of steps taken per day?

```{r, echo=TRUE}
library(ggplot2)
total.steps <- tapply(data$steps, data$date, FUN=sum, na.rm=TRUE)
```

# Histogram of the total number of steps taken each day and 
# mean and median of the total number of steps taken per day

```{r, echo=TRUE}
qplot(total.steps, binwidth=1000, xlab="total number of steps taken each day")
mean(total.steps, na.rm=TRUE)
median(total.steps, na.rm=TRUE)
```

# What is the average daily activity pattern?

```{r, echo=TRUE}
library(ggplot2)
averages <- aggregate(x=list(steps=data$steps), by=list(interval=data$interval),
                      FUN=mean, na.rm=TRUE)
ggplot(data=averages, aes(x=interval, y=steps)) +
  geom_line() +
  xlab("5-minute interval") +
  ylab("average number of steps taken")
```

# Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r, echo=TRUE}
averages[which.max(averages$steps),]
```

# Total number of missing values in the dataset

```{r, echo=TRUE}
missing <- is.na(data$steps)
table(missing)
```

# Replace each missing value with the mean value of its 5-minute interval

```{r, echo=TRUE}
fill.value <- function(steps, interval) {
  filled <- NA
  if (!is.na(steps))
    filled <- c(steps)
  else
    filled <- (averages[averages$interval==interval, "steps"])
  return(filled)
}
filled.data <- data
filled.data$steps <- mapply(fill.value, filled.data$steps, filled.data$interval)
```

# Histogram of the total number of steps taken each day and the mean and median total number of steps taken per day

```{r}
total.steps <- tapply(filled.data$steps, filled.data$date, FUN=sum)
qplot(total.steps, binwidth=1000, xlab="total number of steps taken each day")
mean(total.steps)
median(total.steps)
```

* Imputing missing data resulted in increase in estimates of the total daily number of steps.

# About differences in activity patterns between weekdays and weekends:

* Creating a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r, echo=TRUE}
weekday.or.weekend <- function(date) {
  day <- weekdays(date)
  if (day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"))
    return("weekday")
  else if (day %in% c("Saturday", "Sunday"))
    return("weekend")
  else
    stop("invalid date")
}
filled.data$date <- as.Date(filled.data$date)
filled.data$day <- sapply(filled.data$date, FUN=weekday.or.weekend)
```

* Creating a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

```{r}
averages <- aggregate(steps ~ interval + day, data=filled.data, mean)
ggplot(averages, aes(interval, steps)) + geom_line() + facet_grid(day ~ .) +
  xlab("5-minute interval") + ylab("Number of steps")
```





